Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization

### 2017-12-20 paper/articleApplied.tex
comments

\input{../packageEN}
\input{../packageStyle}

@@

\section{Introduction}
\subsection{Related work}
With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps allow to estimate the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}. These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques. They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day. Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods. %The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}. \ml{je ne vois pas l'intéret de cette dernière phrase}

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels. DYNAMAP project \cite{dynamap_2016} studied the establishment and feasibility of such installations. It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

If sensors networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap. Consequently, the traffic sound level estimation based on measurements is not trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds without distinction between them\cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}. Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level. If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic nous level are presented.

Furthermore, this work and as well as the other works in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context. Although near major roads or ring roads traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels. In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level. In consequence, to be effective on a wide range of sound environments, we propose in this paper to follow the blind source separation paradigm. That is, separating the contribution of the traffic from the other sources within a polyphonic scene.

One of the first and the most widely used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}. The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.  However, ICA is limited to only over determined cases ($N > P$). Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation. A more convenient method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices. It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane. For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded. It has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012} for source separation. They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them. The audio files tested results of a simulation process where a sound background (river or wind) are adding to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing). If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.

@@

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_NMF_EN.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{../image/bloc_diagram_filtrage_EN.pdf}
        \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
		\label{fig:TIR_class_sup}
    \end{subfigure}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{../image/schema_introduction_nmf.pdf}
\caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
\label{fig:example_NMF}
\end{figure}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.8\linewidth]{../image/distanceCosLinDisplay.pdf}
\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
\label{fig:W_ThC_NMF}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{../image/exampleSimScene2.pdf}
\label{fig:exampleSimScene}
\caption{Example of a sound scene composed of 3 sound classes (car, bird, car horn)}
\end{figure}

@@

\begin{figure*}
\centering
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/spectrogramExample.pdf}
   \end{minipage}
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/animals_10-pianoRoll.png}
   \end{minipage}
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{../image/evolutionLpExample.pdf}

@@

The first estimator to determine the traffic sound level is a basic frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz. The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}).

@@

The second estimator is NMF. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.

@@

The dictionary is built from a second sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars (). First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file. $w_t$ is fixed at $wt = [0.5 1]$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra. Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$. A special case is added where the root mean square of \textit{all} the spectrogram is applied. Each audio file generates one element $k$ of $\mathbf{W}$. An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig. \ref{fig:specW}). In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig. \ref{fig:ElementW}).

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{../image/dictionary1.pdf}
        \caption{Sample of 3 seconds of a car passage}
        \label{fig:specW}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{../image/dictionary2.pdf}
        \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
        \label{fig:ElementW}
    \end{subfigure}
\end{figure}

@@

@@

The two first lines resume the error produced by the filter. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguishing the traffic from the others sound sources. Consequently , in low $TIR$ (-12 and -6), where traffic component is discreet, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained. It is then the baseline that will be used to compare the performances of NMF. In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the filter is efficient as it suppress these frequency components whereas for the other sub-classes where low frequencies are present, the error is higher as the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources. The sounds levels are then overestimate (figure \ref{fig:alert_-12}). In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_-12.pdf}
        \caption{}
        \label{fig:alert_-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/LeqTrafficComparison_filter_alert_12.pdf}
        \caption{}
        \label{fig:alert_12}
    \end{subfigure}

@@

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
                \label{fig:TIR_class_sup}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/semi-sup_bar.pdf}
        \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
                \label{fig:TIR_class_semi}
    \end{subfigure}%
    \hfill 
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/THC_bar}
        \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
        \label{fig:TIR_class_thc}
    \end{subfigure}
    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter and each method (supervised, semi-supervised and ThC}
\end{figure*}

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
        \caption{$TIR$ = -12}
        \label{fig:Y_alert-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
        \caption{$TIR$ = 12}
		\label{fig:Y_alert_12}
    \end{subfigure}
    \caption{$\mathbf{W_r}$ for an \textit{alert} scene}
\end{figure*}

Finally, ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}). According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors. On the other hand, on low $TIR$, it does not succeed to be better than semi-supervised NMF.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{../image/dist_-12_12.pdf}    
    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
    \label{fig:dist_-12_12}
\end{figure}

@@

>>>>

\input{packageEN}
\input{packageStyle}

@@

\section{Introduction}
\subsection{Related work}
With the introduction of the European Directive 2002/EC/49, cities over 100 000 inhabitants have to produce road traffic noise maps. These maps depict an estimation of the number of city dwellers exposed to high noise levels and to draw up action plans to reduce it as too long exposures to these noises can generate health problems \cite{who_burden_2017}. These maps are the result of a simulation process based on the estimation of the traffic density on the main roads and the use of sound propagation techniques. They express $L_ {DEN}$ and $L_N$, which are \textit{Day-Evening-Night} and \textit{Night} equivalent A-weighted sound levels respectively. However, these maps introduce lot of uncertainty generated by the numerical tools \cite{van_leeuwen_noise_2015}, by the different calculation methodologies used \cite{leroy_uncertainty_2010}\cite{garg_critical_2014} or even by the calculation procedure of the number of inhabitants exposed to noise \cite{king_implementation_2011}. In addition, the usual road traffic noise maps are static, aggregating the exposure on the two indicators $L_{DEN}$ and $L_N$, thus ignoring the sound levels evolution throughout the day. Since the creation of road traffic noise maps entails long data collection and calculation times, the use of acoustic measurements could facilitate their updating or even the generation of dynamic maps \cite{wei_dynamic_2016}. These measurements can be performed at fixed stations spread all over the cities \cite{Mioduszewski} \cite{mietlicki2012innovative}, which would lead to the availability of the long-term evolution of the traffic noise levels. It can also be performed with  mobile stations \cite{can_exploring_2012} \cite{manvell2004sadmam} covering a larger area with fewer sensors but also sparse time periods. %The clustering between mobile and fixed measurements has been studied in \cite{can_measurement_2014}. \ml{je ne vois pas l'intéret de cette dernière phrase}

Currently, sensor networks in cities are spread for multiple applications (air quality assessment, measurement of meteorological parameters, ...), including the assessment of urban noise levels. DYNAMAP project \cite{dynamap_2016} studied the establishment \ml{c'est anglais ?} and feasibility of such installations. It focuses on sensor installations on specific roads at the city scale in Milan and Rome \cite{bellucci_life_2017}. In a similar way, but reduced to few neighborhoods, the CENSE project\footnote{\url{http://cense.ifsttar.fr/}} \cite{picaut2017characterization} aims to combine \textit{in situ} observations, from a sensor network, and numerical data, from noise modeling, through data assimilation techniques.

If sensors networks could improve road traffic noise estimation compared with simulated maps, the issue of the correct estimation from  measurements of the traffic sound level is still unsolved \cite{Mioduszewski}. Indeed, the urban sound environment is a complex environment gathering lots of different sounds (car passages, voices, bird's whistles, car horn \dots) that can overlap. Consequently, the traffic sound level estimation based on measurements is also not a trivial task.
Many recent works have focused on the detection or recognition tasks of environmental sounds without distinction between them\cite{heittola_sound_2011}, \cite{defreville_automatic_2006}, \cite{dufaux_automatic_2000}, \cite{chu_environmental_2009}. A two step process is generally followed : describe the audio files with a set of features (Spectrum Gravity Spectrum, harmonicity, Mel-Frequency Cepstral Coefficient \dots) and classify them with the help of classifiers (Support Vector Machines, Gaussian Mixture Models, Hidden Markov Model, Artifical Neural Networks). A description of there features and classifiers can be found in \cite{cowling_comparison_2003} and their application can be found in \cite{shen_environmental_2012}, \cite{beritelli_pattern_2008}, \cite{couvreur_automatic_2004}.

Recently, an Anomalous Noise Events Detector has been generated in \cite{socoro_anomalous_2017} to detect the sound sources from labeled recordings that are not related to the traffic component in order not to take them into account on the estimation of the traffic sound level. If the detection of the road traffic noise is good, the detection of these anomalous noise events stay weak and no information on the improvement on the estimation of the traffic  level are presented. Furthermore, this work and as well as the other works in the detection or recognition tasks, do not address the overlap of environmental sounds in an urban context. Although near major roads or ring roads traffic is predominant on all other sound sources, there are many places where road traffic overlaps with other sound sources that contribute significantly to the overall sound levels. In such case, the only detection of the traffic component does not make it possible to determine precisely its noise level.

In consequence, to be effective on a wide range of sound environments, we propose in this paper to follow the blind source separation paradigm. That is, separating the contribution of the traffic from the other sources within a polyphonic scene.

\ml{tu passe beaucoup de temps a taper sur socoro et peu a parler de ton approche. Ce n'est pas une bonne pratique. Je dirais ils font ca, nous aussi on fait ca c'est super, et nous on le fais comme ca, regardez tout ce que ca permet (et implicitement eux ils peuvent pas).}

\ml{tu plonges tres vite dans la technique}

One of the first and the most widely \ml{faux} used techniques to do so is the Independent Component Analysis \cite{comon_independent_1994}. The principle is to decompose $N$ recorded signals to a sum of $P$ independent sound sources weighted by linear relations. This method is most of all suited for the 'cocktail party' issue where one tries to capture a signal among noise.  However, ICA is limited to only over determined cases ($N > P$). Furthermore, if it is suited for indoor environments where the number of sound sources is constant, it can not be fitted for an outdoor environment where the number of sources is unknown and variable and, moreover, it would be necessary to mount multiples sensors on one point to perform the source separation \ml{je ne vois pas ou est le probleme, il y a des gens qui le font}. A more convenient method is Non-negative Matrix Factorization (NMF) \cite{lee_learning_1999} which consists in approximating the magnitude spectrogram of an audio file from the product of two matrices. It has been widely used in the audio domain, \cite{smaragdis_non-negative_2003} \cite{wilson_speech_2008} \cite{mesaros_sound_2015}, and has already been employed for the source separation task of monaural signals of speech and music \cite{wang_musical_2005} \cite{wilson_speech_2008}. By design, this method deals reasonably well with the overlaping sound sources as soon as the overlap can be resolved on the time/frequency plane. For the environmental sounds, the method has been used for the geo-localisation and classification of the sound environment, like in \cite{kumar_audio_2016} where NMF is used to classify the audio files according to the 10 cities where they have been recorded. It has also been used by Innami and Kasai in the unsupervised case \cite{satoshi_innami_nmf-based_2012} for source separation. They proposed a source separation in two steps by separating the sound background from the events first and by separating the events between them. The audio files tested results of a simulation process where a sound background (river or wind) are added to two sound events (school chime, announcement, frog croaking, dog barging and bell ringing).

 If the method proposed is interesting, the main issue here is the small size of the database (only 9 sounds) on which the algorithms are tested while some sounds (frog and river) are not representative of sounds that can be found in cities.
 \ml{encore une fois, utilise la critique implicite. On peut tout critiquer, surtout un papier qui n'est pas encore publié, fais attention.

 faire une liste des contributions avec 1) 2) 3) et finir par un paragraphe de structure}

@@

\ml{tu melange contribution technique et protocole experimental, ca va pas

pas de soussection a l'intro}

@@

\ml{on evite les parentheses en Anglais, pour les ref, faire , see \ref{}}

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_NMF_EN.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=.45\linewidth]{figures/bloc_diagram_filtrage_EN.pdf}
        \caption{Block diagrams summed up the different step of the process for the frequency low-pass filter \ref{fig:blocDiagramFilter} and for NMF \ref{fig:blocDiagramNMF}}
		\label{fig:TIR_class_sup}
    \end{subfigure}
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.9\linewidth]{figures/schema_introduction_nmf.pdf}
\caption{Example of a simple NMF  for urban sound mixture, $\mathbf{W}$ and $\mathbf{V}$ are composed of 3 elements (car passages, car horn and bird's whitles)}
\label{fig:example_NMF}
\end{figure}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.8\linewidth]{figures/distanceCosLinDisplay.pdf}
\caption{Example of the $\mathbf{W}_{traffic}$ extraction from the sorted cosine similarity with a threshold $t = 0.6$. The $82$-nd first elements are considered as traffic component.}
\label{fig:W_ThC_NMF}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{figures/exampleSimScene2.pdf}
\label{fig:exampleSimScene}
\caption{Example of a sound scene composed of 3 sound classes (car, bird, car horn)}
\end{figure}

@@

\begin{figure*}
\centering
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{figures/spectrogramExample.pdf}
   \end{minipage}
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{figures/animals_10-pianoRoll.png}
   \end{minipage}
   \begin{minipage}[c]{.32\linewidth}
      \includegraphics[width =\linewidth]{figures/evolutionLpExample.pdf}

@@

The first estimator to determine the traffic sound level is a frequency low-pass filter which depend only on the cut-off frequencies $f_c$ = [500 1k 2k 5k 10k 20k] Hz. The spectrograms $\mathbf{V}$ are filtered and the remaining energy is then considered as traffic component (eq. \ref{eq:v_tr_filtered}).

@@

The second estimator is the proposed scheme, based on several flavors of the NMF framework. Multiples parameters are involved here between the dictionary learning and the own parameters of the NMF.

@@

The dictionary is built from a second sound database dedicated specifically to this task. It is composed of 53 audio files of passing cars. These records have been made on the Ifsttar's runway with the same conditions that the records made for the \textit{SimScene} database but with two different cars. First, for each audio file, its spectrogram is calculated with fixed parameters ($w$, 50 $\%$ overlap, $nfft$). Then time/frequency windows with $w_t \times F$ dimension are applied without overlapping on the spectrogram in order to consider several spectrum for each audio file. $w_t$ is fixed at $wt = [0.5 1]$ second. In each window, the root mean square value is calculated on each frequency bin to reduce the different spectrum in one spectra. Since the number of elements given by processing all the sound database is high, in order to reduce the computational and delete redundant information, a $K$-means clustering is applied to reduce the number of spectrum to $K = \left[ 25, 50, 100 \right]$. A special case is added where the root mean square of \textit{all} the spectrogram is applied. Each audio file generates one element $k$ of $\mathbf{W}$. An example that illustrates the process can be found on figure \ref{fig:dictionaryExtraction} on 3 seconds extract of a spectrogram of a car passage (fig. \ref{fig:specW}). In the case where $w_t$ = 1 second , 3 elements are therefore extracted of the spectrogram while in the case where $w_t$ = \textit{all}, all the spectrogram is reduced to one element (fig. \ref{fig:ElementW}).

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{figures/dictionary1.pdf}
        \caption{Sample of 3 seconds of a car passage}
        \label{fig:specW}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.47\linewidth}
        \centering
       \includegraphics[width=\linewidth]{figures/dictionary2.pdf}
        \caption{Elements of the dictionary got for $w_t = 1$ second (3 spectrum) and  for $w_t = all$ (1 spectra) }
        \label{fig:ElementW}
    \end{subfigure}
\end{figure}

@@

\ml{replace parameters by experimental factors}

@@

The two first lines show the error produced by the baseline filter. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinguishing the traffic from the others sound sources. Consequently, in low $TIR$ (-12 and -6), where traffic component is scarce, the error is more important than in high $TIR$ (6 and 12) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained. It is then the baseline that will be used to compare the performances of NMF. In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, the filter is efficient as it suppress these frequency components whereas for the other sub-classes where low frequencies are present, the error is higher as the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources. The sounds levels are then overestimate (figure \ref{fig:alert_-12}). In opposite, in high $TIR$, a low-pass filter at $f_c$ = 500 Hz removes too much energy from the traffic which has the consequence to underestimate the sound levels (figure \ref{fig:alert_12}).\\

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_-12.pdf}
        \caption{}
        \label{fig:alert_-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/LeqTrafficComparison_filter_alert_12.pdf}
        \caption{}
        \label{fig:alert_12}
    \end{subfigure}

@@

\ml{remplacer ThC NMF par TINMF pour Thresholded initialized nmf}

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/filter_bar.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/sup_bar.pdf}
        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
                \label{fig:TIR_class_sup}
    \end{subfigure}

    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/semi-sup_bar.pdf}
        \caption{Semi-supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 20 kHz}
                \label{fig:TIR_class_semi}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/THC_bar}
        \caption{ThC NMF, $\beta$ = 2, third octave spectral representation, $f_c$ = 20 kHz}
        \label{fig:TIR_class_thc}
    \end{subfigure}
    \caption{$MAE$ error for each sub-class and $TIR$ according to the the best results with the filter and each method (supervised, semi-supervised and ThC}
\end{figure*}

@@

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_-12_01.pdf}
        \caption{$TIR$ = -12}
        \label{fig:Y_alert-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/Y_alert_TPR_12_01.pdf}
        \caption{$TIR$ = 12}
		\label{fig:Y_alert_12}
    \end{subfigure}
    \caption{$\mathbf{W_r}$ for an \textit{alert} scene}
\end{figure*}

The ThC NMF with a threshold fixed at $t = 0.54$, with a third octave spectral representation and $\beta$ = 2,  offers the lowest average results (Table \ref{tab:results}). According to each $TIR$, with the exception of $TIR$ = 0, all the mean errors are inferior to the baseline and supervised NMF errors. On the other hand, on low $TIR$, it does not succeed to be better than semi-supervised NMF.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{figures/dist_-12_12.pdf}
    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
    \label{fig:dist_-12_12}
\end{figure}

@@

\ml{non, une conclusion conclue, c'est tout.

rappel des resultats experimentaux et conclusion sur l'approche. Rappel des contributions et impact sur la communaute.
Petite ouverture sur le potentiel de l'approche: agnostique en terme de source, peut etre contraint temporellement citation, reste raisonable en terme de cout de calcul. }
